---
title: "Reproducible Research: Peer Assessment 1"
author: "Subramaniam"
date: "January 9, 2016"
output: 
  html_document: 
    keep_md: yes
keep_md: true
---
This is An R Markdown document to demonstrate ease of reproducability.  
  Here we are analyzing Human ACTIVITY data.  
  The activity data is captured at 5 min intervals and has the following fields *"steps", "date",  and "interval"*

The following analysis is incorporated in this document:

  1. Loading and preprocessing the data
  2. What is mean total number of steps taken per day?
  3. What is the average daily activity pattern?
  4. Imputing missing values
  5. Are there differences in activity patterns between weekdays and weekends?
  
###Libraries used  
```{r  warning=FALSE, message=FALSE}
library(readr)
library(lubridate)
library(dplyr)
library(ggvis)
library(lattice)
```

Loading and Preprocessing

```{r echo=TRUE}
activity_asis <- read_csv("~//RSourceData//activity.csv")
activity <- activity_asis[!is.na(activity_asis$steps),]
```

Calculating *mean* total number of steps taken per day
```{r echo=TRUE}
activity %>%
  group_by(date) %>%
  summarise(Total_Number_Steps_PD = sum(steps)) -> Steps_PD

Steps_PD %>% ggvis(~date, ~Total_Number_Steps_PD, fill:="blue") %>% layer_bars() %>% add_axis("x", title = "Date") %>% add_axis("y", title = "Total Number Of Steps Per Day", title_offset = 50)
```

The *mean* and *median* of the total number of steps taken per day
```{r echo=TRUE}
mean(Steps_PD$Total_Number_Steps_PD)
median(Steps_PD$Total_Number_Steps_PD)
```

The average daily activity pattern
```{r echo = TRUE}
activity %>%
  group_by(interval) %>%
  summarise(Average_Across_Days = mean(steps)) -> Average_Steps_Per_Interval

Average_Steps_Per_Interval %>% 
  ggvis(~interval,~Average_Across_Days) %>%
  layer_lines()
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo = TRUE, message=FALSE}
filter(Average_Steps_Per_Interval, Average_Across_Days == max(Average_Steps_Per_Interval$Average_Across_Days))
```

The total number of missing values in the dataset 
```{r echo = TRUE}
activity_asis %>% filter (is.na(steps)) %>% summarise(Missing_Steps_Count = n())
```

Create a new dataset that is equal to the original dataset but with the missing data filled in. Assigning median of Average steps per interval to the missing values based on corresponding intervals
```{r echo = TRUE}
activity_asis[is.na(activity_asis$steps), ]$steps <- median(Average_Steps_Per_Interval$Average_Across_Days)
```

Histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r dispLater, echo = TRUE, results='hide'}

activity_asis %>%
group_by(date) %>%
summarise(Total_Number_Steps_PD = sum(steps)) %>%
ggvis(~date,~Total_Number_Steps_PD,fill:="blue") %>%
layer_bars()%>% add_axis("x", title = "Date") %>% add_axis("y", title = "Total Number Of Steps Per Day (w/ NA Replaced)", title_offset = 50)

activity_asis %>%
group_by(date) %>%
summarise(Total_Number_Steps_PD = sum(steps))%>% summarize(Mean=mean(Total_Number_Steps_PD),Median=median(Total_Number_Steps_PD))
```
```{r ref.label='dispLater', echo=FALSE, message=FALSE}
```
What is the impact of imputing missing data on the estimates of the total daily number of steps?

* Min and Max remains unchanged
* Distribution improves
* Mean and Median change and are less than ones when NA were eleminated.

WeekEnd and WeekDay activity

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo=TRUE, results='hide'}
activity_asis$dayType <- as.factor(ifelse(weekdays(activity_asis$date) %in% c("Sunday","Saturday"),"WeekEnd","WeekDay"))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r echo=TRUE}
activity_asis %>%
  group_by(dayType,interval) %>%
  summarise(Average_Across_DayTypes = mean(steps)) -> activity_dayType

xyplot(Average_Across_DayTypes ~ interval | factor(dayType), data=activity_dayType
       ,layout = c(2, 1), stack = TRUE,pch=c(19,0),
       type = 'l',
       main="Average Number of Steps Taken
       \nAveraged Across All Weekday Days or Weekend Days",
       xlab="5-Minute Interval (military time)",
       ylab="Average Number of Steps Taken")
```

